---
title: "Seurat_v_Scanpy"
output: html_document
date: "2024-01-01"
---

HYPERPARAMETERS
```{r}
# Random seeds
read_downsample_seed <- "0" # random seqtk seed for downsampling reads - 0 for no downsampling, integer >1 for downsampled seed
R_random_seed <- 100 # including for cell downsampling

# Downsample fractions
seu_read_fraction_after_downsampling <- "1.0" # fraction of reads after downsampling - any number from (0,1.0]
scan_read_fraction_after_downsampling <- "1.0" # fraction of reads after downsampling - any number from (0,1.0]

seu_cell_fraction_after_downsampling <- "1.0" # fraction of cells after downsampling - any number from (0,1.0]
scan_cell_fraction_after_downsampling <- "1.0" # fraction of cells after downsampling - any number from (0,1.0]

# Matrix generation methods
seu_matrix_generation_method <- "kb" # str["kb", "cellranger"]
seu_matrix_generation_method_version <- "0.28.0"
seu_matrix_include_unspliced <- FALSE # kb only; if TRUE --> .total.mtx (kb SUA); if FALSE --> .cell.mtx (KB SA) or .mtx (older versions of KB)
seu_matrix_source <- "generated" # "generated" (kb/cellranger count) or "downloaded" (eg 10x website)
seu_matrix_qc <- "raw" # "raw" or "filtered"

scan_matrix_generation_method <- "kb" # str["kb", "cellranger"]
scan_matrix_generation_method_version <- "0.28.0"
scan_matrix_include_unspliced <- FALSE # kb only; if TRUE --> .total.mtx (kb SUA); if FALSE --> .cell.mtx (KB SA) or .mtx (older versions of KB)
scan_matrix_source <- "generated" # "generated" (kb/cellranger count) or "downloaded" (eg 10x website)
scan_matrix_qc <- "raw" # "raw" or "filtered"

# Function argument and input settings
analysis_methods <- "default" # str["default", "seurat_like", "scanpy_like"]
data_input <- "default" # str["default", "seurat", "scanpy"]

# Package versions
seurat_version <- "4.3.0"
scanpy_version <- "1.9.5"

# Custom parameters
inflection_UMI_manual <- NULL # number >=0; or NULL to have automatic selection, especially necessary for lower fracs (e.g., 30 for frac=0.02, 20 for frac=0.01)
min_cells <- 3
min_features <- 200
max_pct_mct <- 20 # default 5
max_n_genes_by_counts_scanpy <- 12000 # default 2500

seu_num_pcs <- 50 # number 1-50; or NULL to select after elbow plot visualization
scan_num_pcs <- 50
umap_knn_k <- 50
umap_leiden_clustering_resolution <- 0.8

# Output path specifications
project_base_path <- "/workspace/analysis"
data_name <- "5k_pbmc_v3"
include_seeds_in_file_paths <- FALSE
save_data <- FALSE
dpi <- 350
```

R Setting up variables
```{r}
set.seed(R_random_seed)
options(max.print = 130)

seu_read_fraction_after_downsampling <- gsub("\\.", "_", as.character(seu_read_fraction_after_downsampling)) # fraction of reads after downsampling, as string representation and using underscores in place of decimal
scan_read_fraction_after_downsampling <- gsub("\\.", "_", as.character(scan_read_fraction_after_downsampling))
seu_cell_fraction_after_downsampling <- gsub("\\.", "_", as.character(seu_cell_fraction_after_downsampling))
scan_cell_fraction_after_downsampling <- gsub("\\.", "_", as.character(scan_cell_fraction_after_downsampling))

seu_matrix_generation_method_version <- gsub("\\.", "_", seu_matrix_generation_method_version)
scan_matrix_generation_method_version <- gsub("\\.", "_", scan_matrix_generation_method_version)

seu_included_transcripts <- ifelse(seu_matrix_include_unspliced == TRUE, "sua", "sa")
scan_included_transcripts <- ifelse(scan_matrix_include_unspliced == TRUE, "sua", "sa")

seu_matrix_generation_method_full <- glue::glue("{seu_matrix_generation_method}{seu_matrix_generation_method_version}_{seu_matrix_qc}_{seu_matrix_source}")
scan_matrix_generation_method_full <- glue::glue("{scan_matrix_generation_method}{scan_matrix_generation_method_version}_{scan_matrix_qc}_{scan_matrix_source}")

# Seurat vs Scanpy
group1_color <- "#D55E00"
group2_color <- "#56B4E9"
```

Data paths
```{r}
seu_data_path <- glue::glue("{project_base_path}/count_matrix_collection/{data_name}/{seu_matrix_generation_method_full}/frac{seu_read_fraction_after_downsampling}_seed{read_downsample_seed}")
scan_data_path <- glue::glue("{project_base_path}/count_matrix_collection/{data_name}/{scan_matrix_generation_method_full}/frac{scan_read_fraction_after_downsampling}_seed{read_downsample_seed}")
```

File path definitions
```{r}
if (seu_matrix_generation_method == scan_matrix_generation_method &&
    seu_matrix_generation_method_version == scan_matrix_generation_method_version &&
    seu_included_transcripts == scan_included_transcripts &&
    seu_matrix_source == scan_matrix_source &&
    seu_matrix_qc == scan_matrix_qc) {
    matrix_generation_method_full <- glue::glue("{seu_matrix_generation_method_full}_{seu_included_transcripts}")
} else {
    matrix_generation_method_full <- glue::glue("seu_{seu_matrix_generation_method_full}_{seu_included_transcripts}__scan_{scan_matrix_generation_method_full}_{scan_included_transcripts}")
}

read_fraction_after_downsampling <- ifelse(seu_read_fraction_after_downsampling == scan_read_fraction_after_downsampling, seu_read_fraction_after_downsampling,
    paste("seu", seu_read_fraction_after_downsampling, "vs", "scan", scan_read_fraction_after_downsampling, sep = "_")
)

cell_fraction_after_downsampling <- ifelse(seu_cell_fraction_after_downsampling == scan_cell_fraction_after_downsampling, seu_cell_fraction_after_downsampling,
    paste("seu", seu_cell_fraction_after_downsampling, "vs", "scan", scan_cell_fraction_after_downsampling, sep = "_")
)

if (include_seeds_in_file_paths) {
    read_fraction_after_downsampling <- glue::glue("{read_fraction_after_downsampling}_seed{read_downsample_seed}")
    cell_fraction_after_downsampling <- glue::glue("{cell_fraction_after_downsampling}_seed{R_random_seed}")
}

output_base_path <- glue::glue("{project_base_path}/output/{data_name}/seuratv{seurat_version}_vs_scanpyv{scanpy_version}/methods_{analysis_methods}/input_{data_input}/{matrix_generation_method_full}/cell_fraction_{cell_fraction_after_downsampling}/read_fraction_{read_fraction_after_downsampling}")

output_data_file_paths <- list(
    markers = glue::glue("{output_base_path}/data_files/markers.rds"),
    results_scan = glue::glue("{output_base_path}/data_files/results_scan.rds"),
    markers2 = glue::glue("{output_base_path}/data_files/markers2.rds"),
    seu_object = glue::glue("{output_base_path}/data_files/seu.rds"),
    scan_adata = glue::glue("{output_base_path}/data_files/adata.h5ad")
)

# FALSE to have no save
file_paths <- list(
    euler_stats_before_QC_file = FALSE, # glue::glue("{output_base_path}/stats/euler_stats_beforeQC.txt"),
    euler_stats_after_QC_file = glue::glue("{output_base_path}/stats/euler_stats_afterQC.txt"),
    pca_knn_clustering_umap_file = glue::glue("{output_base_path}/stats/pca_knn_clustering_umap_stats.txt"),
    de_stats_file = glue::glue("{output_base_path}/stats/de_stats.txt"),
    pre_filtering_upset_cell = FALSE, # glue::glue("{output_base_path}/plots/pre_filtering_upset_cell.tiff"),
    pre_filtering_upset_gene = FALSE, # glue::glue("{output_base_path}/plots/pre_filtering_upset_gene.tiff"),

    knee_plot = FALSE, # glue::glue("{output_base_path}/plots/knee_plot.tiff"),
    umi_scatterplot = FALSE, # glue::glue("{output_base_path}/plots/umi_scatterplot.tiff"),

    violin_counts_comparison <- FALSE, # glue::glue("{output_base_path}/plots/violin_counts_comparison.tiff"),
    seu_violin_file_path = FALSE, # glue::glue("{output_base_path}/plots/seu_violin_plot.tiff"),
    scan_violin_file_path_genes = FALSE, # glue::glue("{output_base_path}/plots/scan_violin_plot_genes.tiff"),
    scan_violin_file_path_counts = FALSE, # glue::glue("{output_base_path}/plots/scan_violin_plot_counts.tiff"),
    scan_violin_file_path_mt = FALSE, # glue::glue("{output_base_path}/plots/scan_violin_plot_mt.tiff"),

    upset_cells = glue::glue("{output_base_path}/plots/upset_cells.tiff"),
    upset_genes = glue::glue("{output_base_path}/plots/upset_genes.tiff"),
    upset_hvgs = glue::glue("{output_base_path}/plots/upset_hvgs.tiff"),
    upset_markers_genes_only = glue::glue("{output_base_path}/plots/upset_marker_genes_only.tiff"),
    upset_markers = glue::glue("{output_base_path}/plots/upset_markers.tiff"),
    euler_before_qc_cell_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_cells_beforeQC.tiff"),
    euler_before_qc_gene_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_genes_beforeQC.tiff"),

    euler_after_qc_cell_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_cells_afterQC.tiff"),
    euler_after_qc_gene_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_genes_afterQC.tiff"),
    euler_after_qc_hvg_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_hvgs_afterQC.tiff"),
    euler_after_qc_marker_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_markers.tiff"),
    euler_after_qc_marker_manual_bonferroni_file_path = FALSE, # glue::glue("{output_base_path}/plots/euler_markers_manual_bonferroni.tiff"),
    euler_after_qc_marker_genes_only = FALSE, # glue::glue("{output_base_path}/plots/euler_markers_genes.tiff"),

    pca_elbow_filepath_combined = FALSE, # glue::glue("{output_base_path}/plots/pca_elbow_combined.tiff"),
    pca_12_overlay_filepath = glue::glue("{output_base_path}/plots/pca_scatterplot_12.tiff"),
    pca_34_overlay_filepath = FALSE, # glue::glue("{output_base_path}/plots/pca_scatterplot_34.tiff"),
    pca_loading_diffs = FALSE, # glue::glue("{output_base_path}/plots/pc_loading_diffs.tiff"),
    pca_eigs_diff = FALSE, # glue::glue("{output_base_path}/plots/pc_eig_diff.tiff"),
    pca_cluster_filepath_seu = FALSE, # glue::glue("{output_base_path}/plots/pca_scatterplot_clusters_seu.tiff"),
    pca_cluster_filepath_scan = FALSE, # glue::glue("{output_base_path}/plots/pca_scatterplot_clusters_scan.tiff"),
    combined_pc_variance_loadings_plot = glue::glue("{output_base_path}/plots/combined_pc_variance_loadings_plot.tiff"),
    jaccards = FALSE, # glue::glue("{output_base_path}/plots/jaccards.tiff"),
    knn_scatterplot = FALSE, # glue::glue("{output_base_path}/plots/knn_scatterplot.tiff"),
    jaccard_degree_scatterplot = glue::glue("{output_base_path}/plots/jaccard_degree_scatterplot.tiff"),
    pheatmap = FALSE, # glue::glue("{output_base_path}/plots/cluster_pheatmap.tiff"),
    alluvial = glue::glue("{output_base_path}/plots/cluster_alluvial.tiff"),
    alluvial_legend = glue::glue("{output_base_path}/plots/cluster_alluvial_legend.tiff"),
    alluvial_legend_high_alpha = glue::glue("{output_base_path}/plots/cluster_alluvial_legend_high_alpha.tiff"),
    umap_seu = glue::glue("{output_base_path}/plots/umap_seu.tiff"),
    umap_scan = glue::glue("{output_base_path}/plots/umap_scan.tiff"),
    umap_seu_clusters_scan = glue::glue("{output_base_path}/plots/umap_seu_clusters_scan.tiff"),
    umap_scan_clusters_seu = glue::glue("{output_base_path}/plots/umap_scan_clusters_seu.tiff"),
    umap_jaccard_degree_scatterplot = glue::glue("{output_base_path}/plots/umap_jaccard_degree_scatterplot.tiff"),
    umap_jaccard_knn_density = glue::glue("{output_base_path}/plots/umap_jaccard_knn_density.tiff"),
    umap_jaccard_knn_density_seu_facet = glue::glue("{output_base_path}/plots/umap_jaccard_knn_density_seu_facet.tiff"),
    umap_jaccard_knn_density_scan_facet = glue::glue("{output_base_path}/plots/umap_jaccard_knn_density_scan_facet.tiff"),
    umap_alluvial = glue::glue("{output_base_path}/plots/umap_alluvial.tiff"),
    umap_alluvial_legend = glue::glue("{output_base_path}/plots/umap_alluvial_legend.tiff"),
    umap_umap_leiden_seu = glue::glue("{output_base_path}/plots/umap_umap_leiden_seu.tiff"),
    umap_umap_leiden_scan = glue::glue("{output_base_path}/plots/umap_umap_leiden_scan.tiff"),
    logFC_histogram_magnitude_file_path = FALSE, # glue::glue("{output_base_path}/plots/logFC_histogram_magnitude.tiff"),
    logFC_histogram_signed_file_path = FALSE, # glue::glue("{output_base_path}/plots/logFC_histogram_signed.tiff"),
    wilcoxon_histogram_magnitude_file_path = FALSE, # glue::glue("{output_base_path}/plots/wilcoxon_histogram_magnitude.tiff"),
    wilcoxon_histogram_signed_file_path = FALSE, # glue::glue("{output_base_path}/plots/wilcoxon_histogram_signed.tiff"),

    logFC_scatterplot_file_path = glue::glue("{output_base_path}/plots/logFC_scatterplot.tiff"),
    wilcoxon_scatterplot_file_path = glue::glue("{output_base_path}/plots/wilcoxon_scatterplot.tiff"),
    logFC_scatterplot_file_path_with_legend = glue::glue("{output_base_path}/plots/logFC_scatterplot_with_legend.tiff"),
    wilcoxon_scatterplot_file_path_with_legend = glue::glue("{output_base_path}/plots/wilcoxon_scatterplot_with_legend.tiff"),
    logFC_scatterplot_outliers_removed_file_path = FALSE, # glue::glue("{output_base_path}/plots/logFC_scatterplot_no_outliers.tiff"),
    wilcoxon_scatterplot_outliers_removed_file_path = FALSE, # glue::glue("{output_base_path}/plots/wilcoxon_scatterplot_no_outliers.tiff"),

    logFC_boxplot_magnitude_file_path = FALSE, # glue::glue("{output_base_path}/plots/logFC_boxplot_magnitude.tiff"),
    logFC_boxplot_signed_file_path = FALSE, # glue::glue("{output_base_path}/plots/logFC_boxplot_signed.tiff"),
    wilcoxon_boxplot_magnitude_file_path = FALSE, # glue::glue("{output_base_path}/plots/wilcoxon_boxplot_magnitude.tiff"),
    wilcoxon_boxplot_signed_file_path = FALSE, # glue::glue("{output_base_path}/plots/wilcoxon_boxplot_signed.tiff"),

    FC_histogram_magnitude_file_path = FALSE, # glue::glue("{output_base_path}/plots/FC_histogram_magnitude.tiff"),
    FC_histogram_signed_file_path = FALSE # glue::glue("{output_base_path}/plots/FC_histogram_signed.tiff")
)

for (path in output_data_file_paths) {
    dir.create(dirname(path), recursive = TRUE, showWarnings = FALSE)
}

if (save_data) {
    for (path in file_paths) {
        if (is.character(path)) {
            # Extract the directory part of the path
            specific_output_path <- dirname(path)

            # Create the directory if it does not exist
            if (!dir.exists(specific_output_path)) {
                dir.create(specific_output_path, recursive = TRUE, showWarnings = FALSE)
            }
        }
    }
} else {
    for (i in seq_along(file_paths)) {
        file_paths[[i]] <- FALSE
    }
}

for (file in c(file_paths$euler_stats_after_QC_file, file_paths$pca_knn_clustering_umap_file, file_paths$de_stats_file)) {
    if (is.character(file)) {
        sink(file = file, append = FALSE)
        sink()
    }
}
```

R Imports
```{r}
conda_env <- "analysis_env"
Sys.setenv(RETICULATE_PYTHON = paste("/home/rstudio/.conda/envs", conda_env, "bin/python3.9", sep = "/"))
library(reticulate)
use_condaenv(conda_env)
library(Seurat)
library(Matrix)
library(tidyverse)
library(patchwork)
library(eulerr)
library(scattermore)
library(DropletUtils)
library(glue)
library(bluster)
library(ggforce)
library(ggplotify)
library(grid)
library(gtable)
library(ggalluvial)
theme_set(theme_bw())

source(glue("{project_base_path}/scripts/data_analysis_helper.R"))
source(glue("{project_base_path}/scripts/plotting_and_stats.R"))
```

Set arguments for functions
```{r}
if (analysis_methods == "default" || analysis_methods == "scanpy_like") {
    scanpy_hvg_flavor <- "seurat"
    n_top_genes <- NULL
    scanpy_scale_max <- NULL
    scanpy_pca_zero_center <- TRUE
    scan_n_neighbors <- 15
    scanpy_clustering_algorithm <- "leiden"
    scanpy_resolution <- 1
    scanpy_cluster_iters <- -1
    scanpy_umap_min_dist <- 0.5
    scanpy_correction_method <- "benjamini-hochberg"
} else if (analysis_methods == "seurat_like") {
    scanpy_hvg_flavor <- "seurat_v3"
    n_top_genes <- 2000
    scanpy_scale_max <- 10
    scanpy_pca_zero_center <- FALSE
    scan_n_neighbors <- 20
    scanpy_clustering_algorithm <- "louvain"
    scanpy_resolution <- 0.8
    scanpy_cluster_iters <- 10
    scanpy_umap_min_dist <- 0.3
    scanpy_correction_method <- "bonferroni"
} else {
    paste(analysis_methods, "is not a valid input for analysis_methods. Please choose from 'default', 'seurat_like', or 'scanpy_like'.")
}

if (analysis_methods == "default" || analysis_methods == "seurat_like") {
    seurat_hvg_flavor <- "vst"
    seu_mean_cutoff <- c(0.1, 8)
    seu_dispersion_cutoff <- c(1, Inf)
    seu_vars_to_regress <- NULL
    seurat_scale_max <- 10
    seu_n_neighbors <- 20
    seurat_clustering_algorithm <- "louvain"
    seu_resolution <- 0.8
    seu_umap_method <- "uwot"
    seu_umap_min_dist <- 0.3
    seu_umap_metric <- "cosine"
    # correction method = bonferroni
} else if (analysis_methods == "scanpy_like") {
    seurat_hvg_flavor <- "mean.var.plot"
    seu_mean_cutoff <- c(0.0125, 3)
    seu_dispersion_cutoff <- c(0.5, Inf)
    seu_vars_to_regress <- c("nCount_RNA", "pct_mt")
    seurat_scale_max <- Inf
    seu_n_neighbors <- 15
    seurat_clustering_algorithm <- "leiden"
    seu_resolution <- 1
    seu_umap_method <- "umap-learn"
    seu_umap_min_dist <- 0.5
    seu_umap_metric <- "correlation"
    # correction method = benjamini-hochberg
} else {
    paste(analysis_methods, "is not a valid input for analysis_methods. Please choose from 'default', 'seurat_like', or 'scanpy_like'.")
}
```

View reticulate python environment
```{r}
py_config()
```

Python imports and setting up variables
```{python}
import os 
import shutil
import scanpy as sc
import pandas as pd
import matplotlib.pyplot as plt
import numpy as np
import anndata
import hdf5plugin
import pickle
import copy
import kb_python.utils as kb_utils

np.random.seed(int(r.R_random_seed))


seu_read_fraction_after_downsampling = r.seu_read_fraction_after_downsampling
scan_read_fraction_after_downsampling = r.scan_read_fraction_after_downsampling

seu_matrix_generation_method = r.seu_matrix_generation_method
scan_matrix_generation_method = r.scan_matrix_generation_method

scan_data_path = r.scan_data_path

scan_num_pcs = r.scan_num_pcs

data_input = r.data_input

if r.n_top_genes:
    n_top_genes = int(r.n_top_genes)
else:
    n_top_genes = r.n_top_genes
    
def save_scanpy_image(filepath):
    if filepath:
        save_file = os.path.basename(filepath)
    else:
        save_file = None
    return save_file
```

Read the data into Seurat and create knee plot:
```{r}
if (seu_matrix_qc == "raw") {
    if (seu_matrix_generation_method == "kb") {
        res_mat <- read_count_output_modified(seu_data_path, name = "cells_x_genes", unspliced = seu_matrix_include_unspliced, tcc = FALSE)
    } else if (seu_matrix_generation_method == "cellranger") {
        res_mat <- Read10X(seu_data_path, gene.column = 1)
    } else {
        print(seu_matrix_generation_method, "is not a valid input for seu_matrix_generation_method")
    }

    if (seu_cell_fraction_after_downsampling != "1_0") {
        total_cells <- ncol(res_mat)
        numeric_seu_cell_fraction_after_downsampling <- gsub("_", ".", seu_cell_fraction_after_downsampling) %>% as.numeric()
        cells_to_sample <- round(total_cells * numeric_seu_cell_fraction_after_downsampling)
        sampled_cells <- sample(total_cells, cells_to_sample)
        res_mat <- res_mat[, sampled_cells]
    }

    tot_counts <- Matrix::colSums(res_mat)
    bc_rank <- barcodeRanks(res_mat)
}
```

Knee plot
```{r}
knee_plot <- make_knee_plot(bc_rank, save = file_paths$knee_plot)
knee_plot
```

Select threshold for knee plot
```{r}
if (seu_matrix_qc == "raw") {
    if (!is.null(inflection_UMI_manual)) {
        UMI_cutoff <- inflection_UMI_manual
    } else {
        inflection_UMI <- metadata(bc_rank)$inflection
        UMI_cutoff <- inflection_UMI
    }
    rank_at_inflection <- max(bc_rank$rank[bc_rank$total > inflection_UMI])
}
```

Apply filtering from knee plot
```{r}
if (seu_matrix_qc == "raw") {
    res_mat_filtered <- res_mat[, tot_counts > UMI_cutoff]
    res_mat_filtered <- res_mat_filtered[Matrix::rowSums(res_mat_filtered) > 0, ]
    dim(res_mat_filtered)
    seu <- CreateSeuratObject(counts = res_mat_filtered, min.cells = min_cells, min.features = min_features)
} else if (seu_matrix_qc == "filtered") {
    seu <- CreateSeuratObject(Read10X(seu_data_path, gene.column = 1), min.cells = min_cells, min.features = min_features)
} else {
    stop("Invalid value for seu_matrix_qc. It should be either 'raw' or 'filtered'.")
}
```

Read the data into Scanpy and apply basic filtering (knee plot)
```{python}
if r.scan_matrix_qc == "raw":  
    if (scan_matrix_generation_method=="kb" and r.scan_matrix_include_unspliced):
        matrix_file = "cells_x_genes.total.mtx"
        adata = kb_utils.import_matrix_as_anndata(f"{scan_data_path}/{matrix_file}",f"{scan_data_path}/cells_x_genes.barcodes.txt",f"{scan_data_path}/cells_x_genes.genes.txt")
    elif (scan_matrix_generation_method=="kb" and not r.scan_matrix_include_unspliced):
        matrix_file = "cells_x_genes.cell.mtx"
        if not os.path.exists(f"{scan_data_path}/{matrix_file}"):
            # If it doesn't exist, use the alternative .mtx file
            matrix_file = "cells_x_genes.mtx"
        adata = kb_utils.import_matrix_as_anndata(f"{scan_data_path}/{matrix_file}",f"{scan_data_path}/cells_x_genes.barcodes.txt",f"{scan_data_path}/cells_x_genes.genes.txt")
    elif scan_matrix_generation_method=="cellranger":
        adata = sc.read_10x_mtx(scan_data_path, var_names='gene_ids')
    
    adata_unfiltered = adata.copy()
    
    if r.scan_cell_fraction_after_downsampling != "1_0":
        total_cells = adata.n_obs
        numeric_scan_cell_fraction_after_downsampling = float(scan_cell_fraction_after_downsampling.replace("_", "."))
        cells_to_sample = round(total_cells * numeric_scan_cell_fraction_after_downsampling)
        sampled_cells_indices = np.random.choice(total_cells, cells_to_sample, replace=False)
        adata = adata[sampled_cells_indices]
    
    # Apply filtering of knee plot
    sc.pp.filter_cells(adata, min_counts=r.UMI_cutoff)  # r.UMI_cutoff (same cutoff as R default) OR custom number
    sc.pp.filter_genes(adata, min_counts=1)
elif r.scan_matrix_qc == "filtered":
    adata = sc.read_10x_mtx(scan_data_path, var_names='gene_ids')
    adata_unfiltered = adata.copy()
else:
    raise ValueError("Invalid value for scan_matrix_qc. It should be either 'raw' or 'filtered'.")

adata.var_names_make_unique()
sc.pp.filter_cells(adata, min_genes=r.min_features)
sc.pp.filter_genes(adata, min_cells=r.min_cells)
```

UMI scatterplot between the two groups
```{r}
res_mat_py <- t(py$adata_unfiltered$X)
rownames(res_mat_py) <- py$adata_unfiltered$var_names$to_list()
colnames(res_mat_py) <- py$adata_unfiltered$obs_names$to_list()

umi_scatterplot <- make_umi_scatterplot(res_mat1 = res_mat, res_mat2 = res_mat_py, UMI_cutoff1 = UMI_cutoff, UMI_cutoff2 = UMI_cutoff, res_mat1_name = "Seurat", res_mat2_name = "Scanpy", point_density = TRUE, color_points = FALSE, save = file_paths$umi_scatterplot)
umi_scatterplot
```

Find list of mitochondrial Ensembl gene names
```{r}
# # ensembl <- biomaRt::useMart("ensembl", dataset = "hsapiens_gene_ensembl")  # Service may be down on this mirror
# ensembl <- biomaRt::useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", mirror = "useast")  #*
# mt_genes <- biomaRt::getBM(attributes = c('ensembl_gene_id', 'external_gene_name'), filters = 'chromosome_name', values = 'MT', mart = ensembl)  #*

mt_genes <- data.frame(ensembl_gene_id = c("ENSG00000210049", "ENSG00000211459", "ENSG00000210077", "ENSG00000210082", "ENSG00000209082", "ENSG00000198888", "ENSG00000210100", "ENSG00000210107", "ENSG00000210112", "ENSG00000198763", "ENSG00000210117", "ENSG00000210127", "ENSG00000210135", "ENSG00000210140", "ENSG00000210144", "ENSG00000198804", "ENSG00000210151", "ENSG00000210154", "ENSG00000198712", "ENSG00000210156", "ENSG00000228253", "ENSG00000198899", "ENSG00000198938", "ENSG00000210164", "ENSG00000198840", "ENSG00000210174", "ENSG00000212907", "ENSG00000198886", "ENSG00000210176", "ENSG00000210184", "ENSG00000210191", "ENSG00000198786", "ENSG00000198695", "ENSG00000210194", "ENSG00000198727", "ENSG00000210195", "ENSG00000210196"))
```

QC and filter high mito cells out of Seurat, Violin plots
```{r}
assay_gene_names <- rownames(seu[["RNA"]])
assay_gene_names_trimmed <- gsub("\\..*", "", assay_gene_names)
common_genes <- intersect(mt_genes$ensembl_gene_id, assay_gene_names_trimmed)
common_genes_with_version <- assay_gene_names[match(common_genes, assay_gene_names_trimmed)]
seu[["pct_mt"]] <- PercentageFeatureSet(seu, features = common_genes_with_version)

violin_plot <- make_violin_plot(seu, show_points = FALSE, color = group1_color, save = file_paths$seu_violin_file_path)
violin_plot

seu <- subset(seu, pct_mt < max_pct_mct)
```

QC and filter high mito cells out of Scanpy
```{python}
sc_mito_genes = r.mt_genes.ensembl_gene_id.tolist()
adata.var['mt'] = adata.var_names.isin(sc_mito_genes)

sc.pp.calculate_qc_metrics(adata, qc_vars=['mt'], percent_top=None, log1p=False, inplace=True)

# Extracting gene names from adata and trimming any versions (like '.1', '.2', etc.)
adata_gene_names_trimmed = [name.split('.')[0] for name in adata.var_names]

# Finding the intersection of mitochondrial genes with genes in adata
common_genes = list(set(sc_mito_genes) & set(adata_gene_names_trimmed))

# Calculate the percentage of mitochondrial reads for each cell
adata.obs['pct_mt'] = np.sum(
    adata[:, [adata.var_names[i] for i, name in enumerate(adata_gene_names_trimmed) if name in common_genes]].X, 
    axis=1
) / np.sum(adata.X, axis=1) * 100

save_genes = save_scanpy_image(r.file_paths[f"scan_violin_file_path_genes"])
save_counts = save_scanpy_image(r.file_paths[f"scan_violin_file_path_counts"])
save_mt = save_scanpy_image(r.file_paths[f"scan_violin_file_path_mt"])

# sc.pl.violin(adata, ['n_genes_by_counts', 'total_counts', 'pct_mt'], jitter=0.4, multi_panel = True)
sc.pl.violin(adata, ['n_genes_by_counts'], jitter=0.4, size=0, save = save_genes)
sc.pl.violin(adata, ['total_counts'], jitter=0.4, size=0, save = save_counts)
sc.pl.violin(adata, ['pct_mt'], jitter=0.4, size=0, save = save_mt)

if r.file_paths[f"scan_violin_file_path_genes"]:
    shutil.move(f"{os.getcwd()}/{save_genes}", r.file_paths[f"scan_violin_file_path_genes"])
    
if r.file_paths[f"scan_violin_file_path_counts"]:
    shutil.move(f"{os.getcwd()}/{save_counts}", r.file_paths[f"scan_violin_file_path_counts"])
    
if r.file_paths[f"scan_violin_file_path_mt"]:
    shutil.move(f"{os.getcwd()}/{save_mt}", r.file_paths[f"scan_violin_file_path_mt"])


pct_cells_over_threshold_mct = np.mean(adata.obs['pct_mt'] > r.max_pct_mct) * 100
print(f"percentage of genes with %mct > threshold: {pct_cells_over_threshold_mct}")

# Filter out cells where the percentage of mitochondrial reads is > 20%
adata = adata[adata.obs['pct_mt'] < r.max_pct_mct, :]

if r.analysis_methods == "default":
    pct_cells_over_threshold_genes_by_counts = np.mean(adata.obs['n_genes_by_counts'] > r.max_n_genes_by_counts_scanpy) * 100
    print(f"percentage of genes with n_genes_by_count > threshold: {pct_cells_over_threshold_genes_by_counts}")
    adata = adata[adata.obs.n_genes_by_counts < r.max_n_genes_by_counts_scanpy, :]
    
# Filter adata to only include the common cells 
cells_adata = adata.obs_names.tolist()

# Convert the cell lists to sets
cells_adata_set = set(cells_adata)
```

Euler plots of cell, gene overlap after QC
```{r}
euler_cell_afterqc <- make_euler_seurat_vs_scanpy(seu, py$adata, comparison = "Cell", save_plot = file_paths$euler_after_qc_cell_file_path, save_stats = file_paths$euler_stats_after_QC_file)
euler_gene_afterqc <- make_euler_seurat_vs_scanpy(seu, py$adata, comparison = "Gene", save_plot = file_paths$euler_after_qc_gene_file_path, save_stats = file_paths$euler_stats_after_QC_file)

euler_cell_afterqc
euler_gene_afterqc
```

Upset plots of cell, gene overlap after QC
```{r}
upset_cell <- make_upset_seurat_vs_scanpy(seu, py$adata, comparison = "Cell", save = file_paths$upset_cells)
upset_gene <- make_upset_seurat_vs_scanpy(seu, py$adata, comparison = "Gene", save = file_paths$upset_genes)
```

Create cell and gene lists
```{r}
seu_inds <- colnames(seu)
seu_genes <- rownames(seu)

scan_inds <- as.vector(py$adata$obs_names$values)
scan_genes <- as.vector(py$adata$var_names$values)

overlapping_inds <- intersect(seu_inds, scan_inds)
overlapping_genes <- intersect(seu_genes, scan_genes)
```

If data_input == "seurat": Apply Cells and Genes from Seurat to Scanpy
```{python}
if data_input == "seurat":
    adata = adata_unfiltered.copy()
    adata = adata[r.seu_inds,r.seu_genes].copy()
    if r.analysis_methods == "default" or r.analysis_methods == "scanpy_like":
        adata.var['mt'] = adata.var_names.isin(sc_mito_genes)
        # Extracting gene names from adata and trimming any versions (like '.1', '.2', etc.)
        adata_gene_names_trimmed = [name.split('.')[0] for name in adata.var_names]

        # Finding the intersection of mitochondrial genes with genes in adata
        common_genes = list(set(sc_mito_genes) & set(adata_gene_names_trimmed))

        # Calculate the percentage of mitochondrial reads for each cell
        adata.obs['pct_mt'] = np.sum(
            adata[:, [adata.var_names[i] for i, name in enumerate(adata_gene_names_trimmed) if name in common_genes]].X,
            axis=1
        ) / np.sum(adata.X, axis=1) * 100
        sc.pp.calculate_qc_metrics(adata, qc_vars=['mt'], percent_top=None, log1p=False, inplace=True)
```

If data_input == "scanpy": Apply Cells and Genes from Scanpy to Seurat
```{r}
if (data_input == "scanpy") {
    if (seu_matrix_generation_method == "kb") {
        res_mat <- read_count_output_modified(seu_data_path, name = "cells_x_genes", unspliced = seu_matrix_include_unspliced, tcc = FALSE)
    } else if (seu_matrix_generation_method == "cellranger") {
        res_mat <- Read10X(seu_data_path, gene.column = 1)
    }
    res_mat <- res_mat[, tot_counts > UMI_cutoff] # uncomment if wanting faster execution
    res_mat <- res_mat[Matrix::rowSums(res_mat) > 0, ] # uncomment if wanting faster execution
    seu <- CreateSeuratObject(counts = res_mat)
    seu <- subset(seu, cells = scan_inds, features = scan_genes)

    assay_gene_names <- rownames(seu[["RNA"]])
    assay_gene_names_trimmed <- gsub("\\..*", "", assay_gene_names)
    common_genes <- intersect(py$sc_mito_genes, assay_gene_names_trimmed)
    common_genes_with_version <- assay_gene_names[match(common_genes, assay_gene_names_trimmed)]
    seu[["pct_mt"]] <- PercentageFeatureSet(seu, features = common_genes_with_version)
}
```

If data_input is not default, then recompute cell and gene lists
```{r}
if (data_input != "default") {
    seu_inds <- colnames(seu)
    seu_genes <- rownames(seu)

    scan_inds <- as.vector(py$adata$obs_names$values)
    scan_genes <- as.vector(py$adata$var_names$values)

    overlapping_inds <- intersect(seu_inds, scan_inds)
    overlapping_genes <- intersect(seu_genes, scan_genes)

    print(paste0("Cell vectors equal: ", all.equal(seu_inds, scan_inds)))
    print(paste0("Gene vectors equal: ", all.equal(seu_genes, scan_genes)))
}
```

Normalization
```{r}
FeatureScatter(seu, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

seu <- NormalizeData(seu, verbose = FALSE)
```

```{python}
if r.scanpy_hvg_flavor != "seurat_v3":
    sc.pp.normalize_total(adata, target_sum=1e4)
```

```{python}
if r.scanpy_hvg_flavor != "seurat_v3":
    sc.pp.log1p(adata)
```

Check equivalency of normalization methods (assuming identical input)
```{r}
mat_py <- py$adata$X
mat_py <- as(t(mat_py), "CsparseMatrix")
mat_r <- GetAssayData(seu, "RNA", layer = "data")
equal_after_normalization <- all.equal(mat_py@x, mat_r@x)

if (file_paths$euler_stats_after_QC_file != FALSE) {
    sink(file_paths$euler_stats_after_QC_file, split = TRUE, append = TRUE)
}

print(glue("Equal after normalization: {equal_after_normalization}"))

if (file_paths$euler_stats_after_QC_file != FALSE) {
    sink()
}
```

Find HVGs
```{r}
seu <- FindVariableFeatures(seu, verbose = FALSE, selection.method = seurat_hvg_flavor, mean.cutoff = seu_mean_cutoff, dispersion.cutoff = seu_dispersion_cutoff, nfeatures = 2000)

top10 <- head(VariableFeatures(seu), 10)
LabelPoints(VariableFeaturePlot(seu), points = top10, repel = TRUE)
```

```{python}
sc.pp.highly_variable_genes(adata, n_top_genes = n_top_genes, flavor = r.scanpy_hvg_flavor)  # Sometimes this cell must be run manually when scanpy_hvg_flavor == "seurat_v3"

scanpy_highly_variable_genes = adata.var.index[adata.var['highly_variable']]
scanpy_highly_variable_genes_list = adata.var[adata.var['highly_variable']].index.tolist()
```

Euler plot of HVG overlap
```{r}
euler_hvg_afterqc <- make_euler_seurat_vs_scanpy(seu, py$adata, comparison = "HVG", save_plot = file_paths$euler_after_qc_hvg_file_path, save_stats = file_paths$euler_stats_after_QC_file)
euler_hvg_afterqc

upset_hvg <- make_upset_seurat_vs_scanpy(seu, py$adata, comparison = "HVG", save = file_paths$upset_hvgs)
```

```{python}
if r.scanpy_hvg_flavor == "seurat_v3":
    sc.pp.normalize_total(adata, target_sum=1e4)
```

```{python}
if r.scanpy_hvg_flavor == "seurat_v3":
    sc.pp.log1p(adata)
    
adata.raw = adata
```

If data_input == "scanpy": Apply HVGs from Scanpy to Seurat
```{r}
if (data_input == "scanpy") {
    VariableFeatures(seu) <- py$scanpy_highly_variable_genes_list
}

is_hvg_r <- rownames(seu) %in% VariableFeatures(seu)
```

If data_input == "seurat": Apply HVGs from Seurat to Scanpy
```{python}
if data_input == "seurat": 
    adata.var = adata.var.assign(highly_variable = r.is_hvg_r)
    scanpy_highly_variable_genes = adata.var.index[adata.var['highly_variable']]
    scanpy_highly_variable_genes_list = adata.var[adata.var['highly_variable']].index.tolist()
```

Make a combined list of HVGs
```{r}
hvgs <- list(Seurat = VariableFeatures(seu), Scanpy = py$scanpy_highly_variable_genes_list)
```

Scanpy-only: Keep only HVGs, regress out features
```{python}
adata = adata[:, adata.var.highly_variable]

if r.analysis_methods == "default" or r.analysis_methods == "scanpy_like":
    sc.pp.regress_out(adata, ['total_counts', 'pct_mt'])
```

Scaling +/- regression
```{r}
seu <- ScaleData(seu, verbose = FALSE, scale.max = seurat_scale_max, vars.to.regress = seu_vars_to_regress)
```

```{python}
sc.pp.scale(adata, max_value=r.scanpy_scale_max)
```

PCA
```{r}
seu <- RunPCA(seu, npcs = 50, verbose = FALSE)
seu_elbow <- Seurat::ElbowPlot(seu, 50)
seu_elbow

if (is.null(seu_num_pcs)) {
    seu_num_pcs <- 50 # optimize as needed
}
```

```{python}
sc.tl.pca(adata, svd_solver='arpack', zero_center = r.scanpy_pca_zero_center)
sc.pl.pca_variance_ratio(adata, log=True, n_pcs=50)

if scan_num_pcs == None:
    scan_num_pcs = 50  # optimize as needed

if r.analysis_methods == "seurat_like":
    scan_num_pcs = r.seu_num_pcs
```

Scree plot
```{r}
if (analysis_methods == "scanpy_like") {
    seu_num_pcs <- py$scan_num_pcs
}

tot_variance <- Misc(Reductions(seu, "pca"))[["total.variance"]]
var_explained <- Stdev(seu, reduction = "pca")^2 / tot_variance
var_explained_py <- py$adata$uns[["pca"]][["variance_ratio"]]

eigs_df <- tibble(
    Seurat = var_explained,
    Scanpy = var_explained_py,
    PC = 1:50
)
```

```{r}
combined_pc_variance <- plot_var_explained(eigs_df, npcs = 50, save = file_paths$pca_elbow_filepath_combined)
combined_pc_variance
```

PCA scatterplot
```{r}
PCAPlot(seu, cols = group1_color) + theme(legend.position = "none")
```

```{python}
sc.pl.pca(adata)
```

Create a collection of PCA embeddings
```{r}
pca_embeddings <- Embeddings(seu, reduction = "pca")

pca_embeddings_py <- py$adata$obsm["X_pca"]

cell_names_py <- scan_inds

rownames(pca_embeddings_py) <- cell_names_py
colnames(pca_embeddings_py) <- colnames(pca_embeddings)

all.equal(cell_names_py, rownames(pca_embeddings))
```

Overlay PCA scatterplots
```{r}
pca12_plot <- plot_pca_compare(pca_embeddings, pca_embeddings_py, save = file_paths$pca_12_overlay_filepath)
pca34_plot <- plot_pca_compare(pca_embeddings, pca_embeddings_py, pcs = 3:4, save = file_paths$pca_34_overlay_filepath)

pca12_plot
pca34_plot
```

Plot PCA eigenvectors (loadings), eigenvalues
```{r}
is_hvg_py <- py$adata$var$highly_variable

pca_loadings_seu <- Loadings(seu, reduction = "pca")
pca_loadings_py <- py$adata$varm["PCs"]
pca_loadings_py <- pca_loadings_py[is_hvg_py, ]
rownames(pca_loadings_py) <- hvgs$Scanpy
colnames(pca_loadings_py) <- colnames(pca_loadings_seu)

df_loadings <- make_pc_diffs_df(list(
    Seurat = pca_loadings_seu,
    Scanpy = pca_loadings_py
), npcs = 50)

mean_loadings_diff <- mean(df_loadings$differences[1:3])

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink(file_paths$pca_knn_clustering_umap_file, split = TRUE, append = TRUE)
}

print(glue("Mean loading difference of PC1-3: {mean_loadings_diff}"))

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink()
}

mylist <- list(
    Seurat = pca_loadings_seu,
    Scanpy = pca_loadings_py
)

loading_diffs <- plot_loading_diffs(df_loadings, save = file_paths$pca_loading_diffs)

df_eigs <- tibble(
    `Seurat vs. Scanpy` = abs(var_explained - var_explained_py),
    PC = 1:50
)

df_eigs <- df_eigs |>
    pivot_longer(-PC, names_to = "type", values_to = "value")

eigs_diff <- plot_eigs_diffs(df_eigs, save = file_paths$pca_eigs_diff)

loading_diffs
eigs_diff
```

Combine scree plot and eigenvector plot
```{r}
combined_plot <- make_combined_pc_variance_loadings_plot(combined_pc_variance, loading_diffs, save = file_paths$combined_pc_variance_loadings_plot)
combined_plot
```

If data_input == "scanpy": Apply PCs from Scanpy to Seurat
```{r}
if (data_input == "scanpy" && identical(seu_inds, scan_inds)) {
    seu[["pca"]] <- CreateDimReducObject(embeddings = as.matrix(pca_embeddings_py), key = "PC_", assay = DefaultAssay(seu))
}
```

If data_input == "seurat": Apply PCs from Seurat to Scanpy
```{r}
if (data_input == "seurat" && identical(seu_inds, scan_inds)) {
    py$adata$obsm["X_pca"] <- pca_embeddings
}
```

Neighbors
```{r}
seu <- FindNeighbors(seu, reduction = "pca", dims = 1:seu_num_pcs, k.param = seu_n_neighbors)
snn_graph_seu <- seu@graphs$RNA_snn
knn_graph_seu <- seu@graphs$RNA_nn
```

```{python}
sc.pp.neighbors(adata, n_neighbors=int(r.scan_n_neighbors), n_pcs=int(scan_num_pcs))
snn_graph_scan = adata.obsp['connectivities']
knn_graph_scan = adata.obsp['distances']
```

Plot SNN graph jaccard indices (ie similarity of neighborhoods) and degrees (ie size of neighborhoods)
```{r}
seu_snn_b <- snn_graph_seu > 0
sc_nn_b <- py$snn_graph_scan > 0

if (!identical(seu_inds, scan_inds)) {
    seu_snn_b <- seu_snn_b[overlapping_inds, overlapping_inds]
    sc_nn_b <- sc_nn_b[overlapping_inds, overlapping_inds]
}

seu_list <- mat2list(seu_snn_b)
sc_list <- mat2list(sc_nn_b)

jaccards <- find_jaccards(list(Seurat = seu_list, Scanpy = sc_list))

median_jaccard <- median(jaccards$Jaccard)

jaccard_plot <- make_jaccard_plot(jaccards, median_jaccard, save = file_paths$jaccards)

jaccard_plot

nei_sizes <- tibble(
    Seurat = lengths(seu_list),
    Scanpy = lengths(sc_list)
)

nei_pairs <- make_pairwise_df(nei_sizes)

knn_scatterplot <- make_knn_scatterplot(nei_pairs, save = file_paths$knn_scatterplot)

knn_scatterplot

jaccards$degree_ratio <- nei_pairs$value1 / nei_pairs$value2
jaccards$logged_degree_ratio <- log(jaccards$degree_ratio, base = 2)

jaccards$jaccard_logged <- log(jaccards$Jaccard, base = 2)

median_logged_degree_ratio <- median(jaccards$logged_degree_ratio)

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink(file_paths$pca_knn_clustering_umap_file, append = TRUE, split = TRUE)
}

print(glue("Median jaccard of KNN: {median_jaccard}"))
print(glue("Median log degree ratio of KNN: {median_logged_degree_ratio}"))

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink()
}
```

Combine jaccard indices and degree ratios into a single plot
```{r}
jaccard_degree_scatterplot <- make_knn_jaccard_degree_scatterplot(jaccards, save = file_paths$jaccard_degree_scatterplot)
jaccard_degree_scatterplot
```

If data_input == "seurat": Apply KNN and SNN graph from Seurat to Scanpy
```{python}
if data_input == "seurat" and (r.seu_inds == r.scan_inds):
    adata.obsp['connectivities'] = r.snn_graph_seu
    snn_graph_scan = adata.obsp['connectivities']
    
    adata.obsp['distances'] = r.knn_graph_seu
    knn_graph_scan = adata.obsp['distances']
```

If data_input == "scanpy": Apply SNN graph from Scanpy to Seurat
```{r}
if (data_input == "scanpy" && identical(seu_inds, scan_inds)) {
    snn_graph_scan_matrix <- as.matrix(py$snn_graph_scan, rownames.force = NA)
    rownames(snn_graph_scan_matrix) <- rownames(py$adata$obs)
    colnames(snn_graph_scan_matrix) <- rownames(py$adata$obs)
    seu@graphs$RNA_snn <- as.Graph(snn_graph_scan_matrix)
    snn_graph_seu <- seu@graphs$RNA_snn

    knn_graph_scan_matrix <- as.matrix(py$adata$obsp["distances"], rownames.force = NA)
    rownames(knn_graph_scan_matrix) <- rownames(py$adata$obs)
    colnames(knn_graph_scan_matrix) <- rownames(py$adata$obs)
    seu@graphs$RNA_nn <- as.Graph(knn_graph_scan_matrix)
    knn_graph_seu <- seu@graphs$RNA_nn
}
```

Clustering
```{r}
if (seurat_clustering_algorithm == "louvain") {
    seurat_clustering_algorithm <- 1
} else if (seurat_clustering_algorithm == "leiden") {
    seurat_clustering_algorithm <- 4
}

seu <- FindClusters(seu, verbose = FALSE, algorithm = seurat_clustering_algorithm, resolution = seu_resolution)

seu_clusters <- Idents(seu)
```

```{python}
if r.scanpy_clustering_algorithm == "leiden":
    sc.tl.leiden(adata, resolution=r.scanpy_resolution, n_iterations=int(r.scanpy_cluster_iters))
elif r.scanpy_clustering_algorithm == "louvain":
    sc.tl.louvain(adata, resolution=r.scanpy_resolution)
```

PCA scatterplots with clusters
```{r}
new_cluster_pca <- DimPlot(seu, label = TRUE, group.by = "seurat_clusters", label.size = 3) + scale_color_manual(values = ditto_colors) + ggtitle("PCA with clusters")
new_cluster_pca

if (file_paths$pca_cluster_filepath_seu != FALSE) {
    ggsave(file_paths$pca_cluster_filepath_seu, plot = new_cluster_pca, dpi = dpi)
}
```

```{python}
ax = sc.pl.pca(adata, color=r.scanpy_clustering_algorithm, show=False, palette=r.ditto_colors, title="PCA with clusters")

# Retrieve handles and labels for the legend
handles, labels = ax.get_legend_handles_labels()

# Create a new legend that includes all clusters
# You might need to adjust 'ncol' (number of columns) for the best layout
ax.legend(handles, labels, loc='best', ncol=2, fontsize='small')

# Show the plot with the updated legend
plt.show()

if r.pca_cluster_filepath_scan != False:
    plt.savefig(r.pca_cluster_filepath_scan)
```

Compute adjusted Rand index to compare cluster similarity
```{r}
scan_clusters <- py$adata$obs[[scanpy_clustering_algorithm]]
names(scan_clusters) <- scan_inds

if (!identical(seu_inds, scan_inds)) {
    seu_clusters <- seu_clusters[names(seu_clusters) %in% overlapping_inds]
    scan_clusters <- scan_clusters[names(scan_clusters) %in% overlapping_inds]

    cell_order <- names(seu_clusters)
    scan_clusters <- scan_clusters[match(cell_order, names(scan_clusters))]
}

seu_clusters_vector <- as.vector(seu_clusters)
scan_clusters_vector <- as.vector(scan_clusters)
ari_value <- mclust::adjustedRandIndex(seu_clusters_vector, scan_clusters_vector)


if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink(file_paths$pca_knn_clustering_umap_file, append = TRUE, split = TRUE)
}

print(glue("Adjusted Rand index between clusters: {ari_value}"))

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink()
}
```

Heatmap of clusters
```{r}
scan_clusters <- factor(scan_clusters, levels = sort(as.numeric(levels(scan_clusters))))

jacc_seu_sc3 <- linkClustersMatrix(seu_clusters, scan_clusters)

cluster_heatmap <- plot_heatmap(jacc_seu_sc3, ari_value, show_axis_titles = TRUE, show_trees = FALSE, save = file_paths$pheatmap)
cluster_heatmap
```

Alluvial plot of clusters
```{r}
df <- tibble(
    Seurat = seu_clusters,
    Scanpy = scan_clusters
)

clus_df_gather <- get_alluvial_df(df)

clus_df_gather <- clus_df_gather %>% mutate(
    group1_column_original_clusters := as.numeric(as.character(.data[["Seurat"]])),
    group2_column_original_clusters := as.numeric(as.character(.data[["Scanpy"]]))
)

clus_df_gather <- sort_clusters_by_agreement(clus_df_gather, stable_column = "Seurat", reordered_column = "Scanpy")

alluvial_plot <- plot_alluvial(clus_df_gather, color_boxes = TRUE, color_bands = FALSE, save = file_paths$alluvial)
alluvial_plot_legend <- plot_alluvial(clus_df_gather, color_boxes = TRUE, color_bands = TRUE, alluvial_alpha = 0.5, save = file_paths$alluvial_legend)
alluvial_plot_legend_high_alpha <- plot_alluvial(clus_df_gather, color_boxes = TRUE, color_bands = TRUE, alluvial_alpha = 1, save = file_paths$alluvial_legend_high_alpha)

alluvial_plot
alluvial_plot_legend
```

Reorder Scanpy clusters to match ordering in alluvial
```{r}
unique_mapping <- clus_df_gather %>%
    ungroup() %>%
    select(!!sym("Scanpy"), group2_column_original_clusters) %>%
    distinct() %>%
    arrange(group2_column_original_clusters)

scanpy_clusters_df <- data.frame(cell_id = names(scan_clusters), group2_cluster = as.numeric(as.character(scan_clusters)))

mapped_data <- scanpy_clusters_df %>% left_join(unique_mapping, by = c("group2_cluster" = "group2_column_original_clusters"))

mapped_data[["Scanpy"]] <- as.character(mapped_data[["Scanpy"]])

named_vector <- setNames(mapped_data[["Scanpy"]], mapped_data$cell_id)

scanpy_clusters_renumbered <- factor(named_vector)
```

```{r}
seu_cluster_data_original <- Idents(seu)
scan_cluster_data_original <- py$adata$obs[[scanpy_clustering_algorithm]]
```

If data_input == "seurat": Apply cluster data from Seurat to Scanpy
```{python}
if data_input == "seurat" and (r.seu_inds == r.scan_inds):
    adata.obs[r.scanpy_clustering_algorithm] = pd.Categorical(r.seu_clusters)
```

If data_input == "scanpy": Apply cluster data from Scanpy to Seurat
```{r}
if (data_input == "scanpy" && identical(seu_inds, scan_inds)) {
    Idents(seu) <- scan_clusters
    seu$seurat_clusters <- Idents(seu)
}
```

UMAP
```{r}
seu_umap_info <- RunUMAP(seu, dims = 1:seu_num_pcs, min.dist = seu_umap_min_dist, umap.method = seu_umap_method, metric = seu_umap_metric)
# DimPlot(seu_umap_info, reduction = "umap") + scale_color_manual(values = ditto_colors)
```

```{python}
sc.tl.umap(adata, min_dist = r.scanpy_umap_min_dist)
# sc.pl.umap(adata, color=r.scanpy_clustering_algorithm, palette=r.ditto_colors, ax=ax, show=False, title="UMAP (Clustering)")
```

Plot UMAP
```{r}
umap_plots <- plot_umap(group1_umap_info = seu_umap_info$umap@cell.embeddings, group1_clusters = seu$seurat_clusters, group2_umap_info = py$adata$obsm["X_umap"], group2_clusters = scanpy_clusters_renumbered, save = c(file_paths$umap_seu, file_paths$umap_scan))
seu_umap <- umap_plots[[1]]
scan_umap <- umap_plots[[2]]

seu_umap
scan_umap


umap_plots_swapped_clusters <- plot_umap(group1_umap_info = seu_umap_info$umap@cell.embeddings, group1_clusters = scanpy_clusters_renumbered, group2_umap_info = py$adata$obsm["X_umap"], group2_clusters = seu$seurat_clusters, group1_title = "Seurat UMAP with Scanpy clusters", group2_title = "Scanpy UMAP with Seurat clusters", save = c(file_paths$umap_seu_clusters_scan, file_paths$umap_scan_clusters_seu))
seu_umap_scan_clusters <- umap_plots_swapped_clusters[[1]]
scan_umap_seu_clusters <- umap_plots_swapped_clusters[[2]]

seu_umap_scan_clusters
scan_umap_seu_clusters
```

Compute KNN graph of UMAP space
```{r}
seu_umap_data <- seu_umap_info$umap@cell.embeddings
scan_umap_data <- py$adata$obsm["X_umap"]

if (!all.equal(colnames(seu), as.vector(py$adata$obs_names$values))) {
    seu_inds <- colnames(seu)
    scan_inds <- as.vector(py$adata$obs_names$values)

    rownames(scan_umap_data) <- scan_inds
    scan_cluster_data_original <- setNames(scan_cluster_data_original, scan_inds)

    overlapping_inds <- intersect(seu_inds, scan_inds)

    seu_umap_data_filtered <- seu_umap_data[overlapping_inds, ]
    seu_umap_data <- seu_umap_data_filtered[order(rownames(seu_umap_data_filtered)), ]

    scan_umap_data_filtered <- scan_umap_data[overlapping_inds, ]
    scan_umap_data <- scan_umap_data_filtered[order(rownames(scan_umap_data_filtered)), ]

    seu_cluster_data_filtered <- seu_cluster_data_original[overlapping_inds]
    seu_cluster_data_original <- seu_cluster_data_filtered[order(names(seu_cluster_data_filtered))]

    scan_cluster_data_filtered <- scan_cluster_data_original[overlapping_inds]
    scan_cluster_data_original <- scan_cluster_data_filtered[order(names(scan_cluster_data_filtered))]
} else {
    seu_cluster_data_filtered <- seu_cluster_data_original
    scan_cluster_data_filtered <- scan_cluster_data_original
}

seu_umap_knn <- dbscan::kNN(seu_umap_data, k = umap_knn_k)
scan_umap_knn <- dbscan::kNN(scan_umap_data, k = umap_knn_k)
```

Find jaccard indices of KNN graphs from UMAP space
```{r}
jaccards_all_cells <- calculate_knn_jaccards(seu_umap_knn$id, scan_umap_knn$id)

median_jaccard_umap_knn <- median(jaccards_all_cells)

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink(file_paths$pca_knn_clustering_umap_file, append = TRUE, split = TRUE)
}

print(glue("Median jaccard of UMAP KNN: {median_jaccard_umap_knn}"))

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink()
}
```

Plot jaccard indices
```{r}
jaccards_df <- data.frame(Cells = overlapping_inds, JaccardIndex = jaccards_all_cells, seu_clusters = seu_cluster_data_filtered, scan_clusters = scan_cluster_data_filtered)

umap_jaccard_plot <- make_umap_jaccard_plot(jaccards_df, save = file_paths$umap_jaccard_knn_density)
umap_jaccard_plot_seu_facet <- make_umap_jaccard_plot(jaccards_df, facet = "seu_clusters", save = file_paths$umap_jaccard_knn_density_seu_facet)
umap_jaccard_plot_scan_facet <- make_umap_jaccard_plot(jaccards_df, facet = "scan_clusters", save = file_paths$umap_jaccard_knn_density_scan_facet)

umap_jaccard_plot
umap_jaccard_plot_seu_facet
umap_jaccard_plot_scan_facet
```

Run leiden clustering on KNN graphs from UMAP space
```{r}
set.seed(R_random_seed)

seu_umap_knn_clusters <- bluster::clusterRows(seu_umap_data, NNGraphParam(shared = FALSE, k = umap_knn_k, cluster.fun = "leiden", cluster.args = list(resolution_parameter = umap_leiden_clustering_resolution, objective_function = "modularity", n_iterations = 2)))
scan_umap_knn_clusters <- bluster::clusterRows(scan_umap_data, NNGraphParam(shared = FALSE, k = umap_knn_k, cluster.fun = "leiden", cluster.args = list(resolution_parameter = umap_leiden_clustering_resolution, objective_function = "modularity", n_iterations = 2)))

seu_umap_knn_clusters <- reorder_clusters_descending(seu_umap_knn_clusters)
scan_umap_knn_clusters <- reorder_clusters_descending(scan_umap_knn_clusters)
```

Compute ARI and plot alluvial plot of leiden clustering results on KNN graphs from UMAP space
```{r}
ari_value_umap <- mclust::adjustedRandIndex(as.vector(seu_umap_knn_clusters), as.vector(scan_umap_knn_clusters))

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink(file_paths$pca_knn_clustering_umap_file, append = TRUE, split = TRUE)
}

print(glue("Adjusted Rand index between clusters: {ari_value_umap}"))

if (file_paths$pca_knn_clustering_umap_file != FALSE) {
    sink()
}

df_umap <- tibble(
    Seurat = seu_umap_knn_clusters,
    Scanpy = scan_umap_knn_clusters
)

clus_df_gather_umap <- get_alluvial_df(df_umap)

clus_df_gather_umap <- clus_df_gather_umap %>% mutate(
    group1_column_original_clusters := as.numeric(as.character(.data[["Seurat"]])),
    group2_column_original_clusters := as.numeric(as.character(.data[["Scanpy"]]))
)

clus_df_gather_umap <- sort_clusters_by_agreement(clus_df_gather_umap, stable_column = "Seurat", reordered_column = "Scanpy")

umap_alluvial_plot <- plot_alluvial(clus_df_gather_umap, color_boxes = TRUE, color_bands = FALSE, alluvial_alpha = 0.5, save = file_paths$umap_alluvial)
umap_alluvial_plot_legend <- plot_alluvial(clus_df_gather_umap, color_boxes = TRUE, color_bands = TRUE, alluvial_alpha = 0.5, save = file_paths$umap_alluvial_legend)

umap_alluvial_plot_legend
```

Reorder Scanpy UMAP clusters to match ordering in alluvial
```{r}
unique_mapping <- clus_df_gather_umap %>%
    ungroup() %>%
    select(!!sym("Scanpy"), group2_column_original_clusters) %>%
    distinct() %>%
    arrange(group2_column_original_clusters)

scanpy_clusters_df <- data.frame(cell_id = overlapping_inds, group2_cluster = as.numeric(as.character(scan_umap_knn_clusters)))

mapped_data <- scanpy_clusters_df %>% left_join(unique_mapping, by = c("group2_cluster" = "group2_column_original_clusters"))

mapped_data[["Scanpy"]] <- as.character(mapped_data[["Scanpy"]])

named_vector <- setNames(mapped_data[["Scanpy"]], mapped_data$cell_id)

scanpy_clusters_renumbered_umap <- factor(named_vector)
```

UMAP with UMAP Leiden clusters
```{r}
umap_plots <- plot_umap(group1_umap_info = seu_umap_data, group1_clusters = seu_umap_knn_clusters, group2_umap_info = scan_umap_data, group2_clusters = scanpy_clusters_renumbered_umap, group1 = "Seurat", group2 = "Scanpy", save = c(file_paths$umap_umap_leiden_seu, file_paths$umap_umap_leiden_scan))
seu_umap <- umap_plots[[1]]
scan_umap <- umap_plots[[2]]

seu_umap
scan_umap
```

Find markers Seurat
```{r}
if (!file.exists(output_data_file_paths$markers)) {
    if (analysis_methods == "default" || analysis_methods == "seurat_like") {
        markers <- FindAllMarkers(seu)
    } else if (analysis_methods == "scanpy_like") {
        saveRDS(seu, file = output_data_file_paths$seu_object)
        markers <- FindAllMarkers(seu, logfc.threshold = 0, min.pct = 0, return.thresh = 1.0001) # if this crashes R, then run de_alone.R
        markers$p_val_adj <- p.adjust(degs$p_val, method = "BH")
    }
} else {
    markers <- readRDS(output_data_file_paths$markers)
}
```

Find markers Scanpy
```{python}
# adata.obs['leiden'] = adata.obs['leiden'].astype(str)
sc.tl.rank_genes_groups(adata, r.scanpy_clustering_algorithm, use_raw=True, method='wilcoxon', corr_method=r.scanpy_correction_method, pts = True)
# adata.uns['rank_genes_groups']
```

```{r}
result <- get_py_de_results("adata")
```

Scanpy manual filtering if applying seurat-like filtering
```{r}
if (analysis_methods == "seurat_like" && data_input == "seurat") {
    result <- result %>%
        filter(!(pts < 0.1 & pts_rest < 0.1)) %>%
        filter(!(abs(log_fc) < 0.25)) %>%
        filter(p_value < 0.01)

    min_logFC_filtered <- min(abs(result$log_fc), na.rm = TRUE) # log_fc or avg_log2FC - should be 0.25
    max_p_val_filtered <- max(result$p_value, na.rm = TRUE) # p_value or p_val - should be 0.01
    exists_row_filtered <- any(result$pts < 0.1 & result$pts_rest < 0.1) # should be FALSE

    min_logFC_filtered
    max_p_val_filtered
    exists_row_filtered
    cat("Final number of rows in filtered:", nrow(result), "\n")
}
```

Compare marker genes
```{r}
# # Uncomment if running with scanpy_like methods and loading in markers after running de_alone.R
# if (analysis_methods == "scanpy_like") {
#     markers <- readRDS("/workspace/analysis/seu_markers_after_de_scanlike.rds")
# }

seu_filtered_markers <- markers %>%
    filter(p_val_adj < 0.05)

# vectorized_seu_unfiltered_markers <- unique(markers$gene)
vectorized_seu_filtered_markers <- unique(seu_filtered_markers$gene)

scan_filtered_markers <- result %>%
    filter(p_value_adj < 0.05)

vectorized_scan_filtered_markers <- unique(scan_filtered_markers$gene)

markers_euler_genes_only <- make_euler_seurat_vs_scanpy(vectorized_seu_filtered_markers, vectorized_scan_filtered_markers, comparison = "Marker Gene", save_plot = file_paths$euler_after_qc_marker_genes_only, save_stats = file_paths$de_stats_file)
markers_euler_genes_only

upset_marker_gene_only <- make_upset_seurat_vs_scanpy(vectorized_seu_filtered_markers, vectorized_scan_filtered_markers, comparison = "Marker Gene", save = file_paths$upset_markers_genes_only)
```

Stop DE analysis if data input does not align (as without aligned cluster information, DE analysis is not meaningful)
```{r}
if (data_input == "default" || !identical(seu_inds, scan_inds)) {
    if (save_data == TRUE) {
        saveRDS(markers, file = output_data_file_paths$markers)
        saveRDS(result, file = output_data_file_paths$results_scan)
        saveRDS(seu, file = output_data_file_paths$seu_object)
        py$adata$write_h5ad(output_data_file_paths$scan_adata, compression = py$hdf5plugin$FILTERS$zstd)
    }
    sessionInfo()
    stop("data_input == 'default' or cell sets not identical, so not running further DE analysis, which requires clusters to be in agreement.")
}
```

Compare markers
```{r}
# Select gene and cluster columns
seu_markers_df <- markers %>% select(gene = gene, cluster = cluster)
scan_markers_df <- result %>% select(gene = gene, cluster = cluster)

vectorized_seu_markers <- paste(seu_markers_df$gene, seu_markers_df$cluster, sep = "-")
vectorized_scan_markers <- paste(scan_markers_df$gene, scan_markers_df$cluster, sep = "-")

markers_euler <- make_euler_seurat_vs_scanpy(vectorized_seu_markers, vectorized_scan_markers, comparison = "Marker", save_plot = file_paths$euler_after_qc_marker_file_path, save_stats = file_paths$de_stats_file)
markers_euler

upset_markers_all <- make_upset_seurat_vs_scanpy(vectorized_seu_markers, vectorized_scan_markers, comparison = "Marker", save = file_paths$upset_markers)
```

Combine all DE data in one dataframe markers2
```{r}
markers2 <- markers |>
    left_join(result, by = c("cluster", "gene"))

markers2 <- markers2 |>
    dplyr::rename(
        p_val_r = p_val, logFC_r = avg_log2FC, p_val_adj_r = p_val_adj,
        p_val_py = p_value, p_val_adj_py = p_value_adj,
        logFC_py = log_fc
    )

markers2 <- markers2 |>
    mutate(cluster = factor(cluster, levels = as.character(seq_len(length(unique(cluster))) - 1)))

markers2 <- markers2 |>
    group_by(cluster) |>
    mutate(rank_r = seq_along(gene))

markers2$FC_r <- 2^markers2$logFC_r
markers2$FC_py <- 2^markers2$logFC_py
```

Calculate mean magnitude of difference in log fold change between the 2 packages
```{r}
markers2 <- calculate_de_stats(markers2, save = file_paths$de_stats_file)
```

Plot histograms
```{r}
seurat_vs_scanpy_logFC_histogram_magnitude <- plot_differences_histogram_seurat_vs_scanpy(markers2, "logFC_difference_magnitude", title = glue("Histogram of logFC (|Seu-Scan|)"), x_label = "logFC (|Seu-Scan|)", median_or_variance = "median", save = file_paths$logFC_histogram_magnitude_file_path)
seurat_vs_scanpy_logFC_histogram_signed <- plot_differences_histogram_seurat_vs_scanpy(markers2, "logFC_difference_signed", title = glue("Histogram of logFC (Seu-Scan)"), x_label = "logFC (Seu-Scan)", median_or_variance = "variance", save = file_paths$logFC_histogram_signed_file_path)

seurat_vs_scanpy_pvaladj_histogram_magnitude <- plot_differences_histogram_seurat_vs_scanpy(markers2, "pvaladj_difference_magnitude", title = glue("Histogram of p_val_adj (|Seu-Scan|)"), x_label = "p_val_adj (|Seu-Scan|)", median_or_variance = "variance", save = file_paths$wilcoxon_histogram_magnitude_file_path)
seurat_vs_scanpy_pvaladj_histogram_signed <- plot_differences_histogram_seurat_vs_scanpy(markers2, "pvaladj_difference_signed", title = glue("Histogram of p_val_adj (Seu-Scan)"), x_label = "p_val_adj (Seu-Scan)", median_or_variance = "variance", save = file_paths$wilcoxon_histogram_signed_file_path)


seurat_vs_scanpy_logFC_histogram_magnitude
seurat_vs_scanpy_logFC_histogram_signed

seurat_vs_scanpy_pvaladj_histogram_magnitude
seurat_vs_scanpy_pvaladj_histogram_signed
```

Heatmap
```{r}
# Plot heatmap
markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(seu, features = top10$gene, size = 4) + NoLegend() + theme(text = element_text(size = 4.5))
```

Plot scatterplots
```{r}
markers2$p_val_adj_r[markers2$p_val_adj_r == 0] <- .Machine$double.xmin
markers2$p_val_adj_py[markers2$p_val_adj_py == 0] <- .Machine$double.xmin

seurat_vs_scanpy_logFC_scatterplot <- plot_scatterplot_de_logfc(markers2, save = file_paths$logFC_scatterplot_file_path, outliers_excluded = FALSE)
seurat_vs_scanpy_pvaladj_scatterplot <- plot_scatterplot_de_wilcoxon(markers2, save = file_paths$wilcoxon_scatterplot_file_path, outliers_excluded = FALSE)

seurat_vs_scanpy_logFC_scatterplot_with_legend <- plot_scatterplot_de_logfc(markers2, save = file_paths$logFC_scatterplot_file_path_with_legend, outliers_excluded = FALSE, show_legend = TRUE)

seurat_vs_scanpy_logFC_scatterplot
seurat_vs_scanpy_logFC_scatterplot_with_legend

seurat_vs_scanpy_pvaladj_scatterplot
```

Save markers df, seu and scan objects
```{r}
subset_markers2 <- markers2[, c("gene", "cluster", "logFC_py", "logFC_r", "p_val_adj_r", "p_val_adj_py", "logFC_difference_magnitude", "logFC_difference_signed", "pvaladj_difference_magnitude", "pvaladj_difference_signed")]

if (save_data == TRUE) {
    saveRDS(markers, file = output_data_file_paths$markers)
    saveRDS(result, file = output_data_file_paths$results_scan)
    saveRDS(subset_markers2, file = output_data_file_paths$markers2)
    saveRDS(seu, file = output_data_file_paths$seu_object)
    py$adata$write_h5ad(output_data_file_paths$scan_adata, compression = py$hdf5plugin$FILTERS$zstd)
}
```

```{r}
sessionInfo()
```
